---
title: "Multi-omic COVID-19 Manuscript figures"
output:
  html_document:
    df_print: paged
    toc: TRUE
    theme: united
date: "June 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(dplyr)
library(gridExtra)
library(pheatmap)
library(plyr)
library(data.table)
library(openxlsx)
library(cowplot)
library(stringr)
library(corrplot)
library(glmnet)
library(ggfortify)
library(RColorBrewer)
library(randomForest)
library(Rmisc)
library(Rtsne)
library(corrr)
library(pROC)
library(caret)
library(tidyr)
library(psych)
library(mclust)
library(GGally)
library(formattable)
library(car)
library(rstatix)
library(mclust)
library(fossil)
library(foreach)
library(cowplot)
library(grid)
library(tibble)
library(MVN)
library(npmv)
library(wesanderson)
library(scales)
library(lmerTest)
library(see)
library(performance)
library(glmmTMB)
library(DHARMa)
library(ggthemes)
library(effects)
library(caret)
library(scales)
library(survival)
library(survminer)
library(openxlsx)
library(MOFA2)
library(ggExtra)

pal <- wes_palette("Zissou1", 100, type = "continuous") # Heatmap

`%out%` <- function(a,b) ! a %in% b # Need to specify my own operator

cor.mtest <- function(mat, ...) {
  # mat : is a matrix of data
  # ... : further arguments to pass to the native R cor.test function
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat<- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", alternative = "two.sided") # I checked the results of this test against another independent tool and I got the same p-values.
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
} # Get the p-value matrix



my_fn <- function(data, mapping, method="lm", ...){ # for ggpairs
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point(size = 1) 
    #geom_smooth(method=method, ..., level = 0)
  p
}








```




```{r read-n-wrangle,  echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
# Generate penalized expression matrix
unstim <- read.xlsx("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Penalization matrix_CyTOF.xlsx", sheet = 1) %>%
  mutate(stimulation = "Unstim")
LPS <- read.xlsx("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Penalization matrix_CyTOF.xlsx", sheet = 2) %>%
  mutate(stimulation = "LPSCI")
IFN <- read.xlsx("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Penalization matrix_CyTOF.xlsx", sheet = 3) %>%
  mutate(stimulation = "IFNIL")
PI <- read.xlsx("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Penalization matrix_CyTOF.xlsx", sheet = 4) %>%
  dplyr::select(-X22) %>%
  mutate(stimulation = "PI")
pen.df <- rbind(unstim, LPS, IFN, PI) %>%
  melt(id.vars = c("population", "stimulation"), variable.name = "reagent", value.name = "penalty") %>%
  mutate(population = case_when(population == "Ki67CD8T" ~ "Ki67CD8",
                                TRUE ~ population))

cytof.df <- read.csv("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/083120_COVID_StanfordPBMC_all_preprocessed.csv") %>%
  full_join(., pen.df, by = c("population", "reagent", "stimulation")) %>%
  mutate(penalty = case_when(reagent == "frequency" ~ 1,
                             TRUE ~ penalty)) %>%
  mutate(penalized = feature * penalty) %>%
  na.omit() # This deals with the duplicated subjects issue
  #dplyr::filter(sampleID %in% c("007-0011", "55689-0112")) # These two subjects have duplicated data

meta.df <- read.csv("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Clinical_12NOV.csv") %>%
  mutate(sampleID = as.character(sampleID))

olink.df <- read.csv("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/20201029_Gaudilliere_NPX_2020-07-20_corrected.csv") %>%
  mutate(group = mapvalues(sampleID, from = meta.df$sampleID, to = as.character(meta.df$Severity))) %>%
  group_by(sampleID, Assay, group) %>% # TNF, IL6 and IL8 are in all 4 subplexes. Take the
  dplyr::summarize(NPX = mean(NPX)) %>%
  dplyr::rename(sample = sampleID, 
                value = NPX,
                feature = Assay) %>% 
  mutate(view = "olink") %>%
  dplyr::select(sample, feature, view, value, group) %>%
  ungroup
```


```{r mofa prep,  echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
# MOFA prep
cytof.df <- cytof.df %>% 
  mutate(stimulation = case_when(reagent == "frequency" ~ "frequency",
                                 TRUE ~ stimulation)) %>%
  mutate(feature = paste0(population, ".", reagent)) %>%
  dplyr::rename(sample = sampleID,
                value = penalized,
                view = stimulation) %>%
  dplyr::select(sample, feature, view, value, group) %>%
  split(.$view) 

dat <- rbind(cytof.df$frequency, cytof.df$Unstim, cytof.df$IFNIL, cytof.df$LPSCI, cytof.df$PI, olink.df) %>%
  dplyr::select(-group)

MOFAobject <- create_mofa(dat)


# Settings for MOFA calculation
data_opts <- get_default_data_options(MOFAobject)
data_opts$scale_views <- FALSE
model_opts <- get_default_model_options(MOFAobject)
model_opts$num_factors <- 25 # the maximum of 25 factors allowed for each run; however for each initialization, the total number of factors that explain the most variation of the data are calculated (this number is always less than 25)
train_opts <- get_default_training_options(MOFAobject)
train_opts$convergence_mode <- "slow"
train_opts$drop_factor_threshold <- 0.02
train_opts$seed <- 42
train_opts$maxiter <- 10000
```


```{r mofa inits,  echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
n_inits <- 10
MOFAlist <- lapply(seq_len(n_inits), function(it) {
  
  train_opts$seed <- 41 + it
  
  MOFAobject <- prepare_mofa(
    object = MOFAobject,
    data_options = data_opts,
    model_options = model_opts,
    training_options = train_opts
  )
  
  run_mofa(MOFAobject, use_basilisk = TRUE)
})
#saveRDS(MOFAlist, file = "/Users/joshgillard/Desktop/Bioinformatics/Multiomic COVID19/MOFA_COVID.rds")
```


```{r mofa modelselection,  echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
MOFAlist <- readRDS(file = "/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/MOFA_COVID.rds")
MOFAobject.trained <- select_model(MOFAlist)
mofa_fac <- get_factors(MOFAobject.trained)
var.explained <- calculate_variance_explained(MOFAobject.trained)
```


## Model selection {.tabset}

### No. factors and ELBO score

```{r elbo and factors, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 9, fig.width = 9}
i <- 1
res <- data.frame(elbo = numeric(),
                  fac = numeric(),
                  model =  numeric())
for (i in 1:length(MOFAlist)) {
  mofa.tmp <- MOFAlist[[i]]
  elbo.tmp <- max(mofa.tmp@training_stats$elbo)
  fac.tmp <- min(mofa.tmp@training_stats$number_factors)
  res[i, 1] <- elbo.tmp
  res[i, 2] <- fac.tmp
  res[i, 3] <- i
}
MOFA.selection <- res %>%
  ggplot(aes(x = model, y = fac, fill = elbo)) +
  geom_bar(stat = "identity", 
           color = "black", width = 0.9) +
  theme_minimal() +
  scale_fill_gradientn(colors = c("gray97", "darkblue"), 
                       
                       values = scales::rescale(c(0, 1, 32, 38, 40)),
                       guide = "colorbar") +
  labs(y = "number of active factors",
       x = "model trials") +
  scale_x_continuous(breaks = 1:10) +
   theme(axis.title.x = element_text(size = 6.5),
        axis.title.y = element_text(size = 6.5),
        axis.text.x = element_text(angle = 0, size  = 6.5),
        axis.text.y = element_text(size = 6.5),
        plot.title = element_text(size = 6.5),
        legend.text = element_text(size = 6.5),
        legend.title = element_text(size = 6.5),
        strip.text.x = element_text(size = 6.5))
MOFA.selection


# ggsave(plot = MOFA.selection,
#        file = paste("MOFA.selection", ".pdf", sep=""),
#        path = "/Users/joshgillard/Desktop/",
#        width = 6,
#        height = 6,
#        units = "cm",
#        device = "pdf")
```

### Compare factors

In downstream analyses we discover factor 3 and factor 10 of model 6 are correlated with COVID-19 severity. 

in this plot, we can see if those factors are consistently discovered across different random initializations of MOFA

```{r POS n, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 12, fig.width = 20}
compare_factors(MOFAlist)

```


## Model Overview {.tabset}

### Factor correlation

```{r factor corr, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 6, fig.width = 6}
plot_factor_cor(MOFAobject.trained)
# This is a check on the quality of the MOFA factor calculation; ideally, each mofa factor will not be highly correlated with any other mofa factor.
```

### % Var explained

```{r total var explained, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 6, fig.width = 6}
# How much of the joint variation in the data is explained by each dataset?
var.explained %>%
  magrittr::extract2(1) %>%
  unname() %>%
  as.data.frame() %>%
  dplyr::rename(value = 1) %>%
  rownames_to_column("dataset") %>%
  ggplot(aes(x = dataset, y = value)) +
  geom_bar(stat = "identity", fill = "deepskyblue4", 
           color = "black", width = 0.9) +
  theme_minimal() +
  labs(y = "% variance explained",
       x = "")
```


### Factor breakdown

```{r factor var explained, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 6, fig.width = 6}
original <- var.explained %>%
  magrittr::extract2(2) %>%
  unname() %>%
  data.frame() %>%
  rownames_to_column("factors") %>%
  melt(., id.vars = "factors") %>%

  mutate(factors = factor(factors, levels = unique(.$factors))) %>%

  ggplot(aes(x = factors, y = variable)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = paste0(round(value, 2))), size  =  6.5 / (14/5)) +
  coord_equal() +
  theme_minimal() +
  scale_fill_gradientn(colors = c("gray97", "darkblue"), 
                       guide = "colorbar") +
  theme(axis.text.x = element_text(angle = 90))+
  labs(y = "",
       x = "",
       fill = "% variance explained")
# ggsave(plot = original,
#        file = paste("original", ".pdf", sep=""),
#        path = "/Users/joshgillard/Desktop/",
#        width = 18,
#        height = 6,
#        units = "cm",
#        device = "pdf")


mofa.loading <- var.explained %>%
  magrittr::extract2(2) %>%
  unname() %>%
  data.frame() %>%
  rownames_to_column("factors") %>%
  melt(., id.vars = "factors") %>%
  
  mutate(factors = factor(factors, levels = unique(.$factors))) %>%
  
  ggplot(aes(x = factors, y = variable)) +
  geom_tile(aes(fill = value)) +
  #geom_text(aes(label = paste0(round(value, 2))), size  =  6.5 / (14/5)) +
  coord_equal()+
  theme_minimal() +
  scale_fill_gradientn(colors = c("gray97", "darkblue"), 
                       values = scales::rescale(c(0, 1, 32, 38, 40)),
                       guide = "colorbar") +
  labs(y = "",
       x = "",
       fill = "% variance explained") +
  theme(axis.title.x = element_text(size = 6.5),
        axis.title.y = element_text(size = 6.5),
        axis.text.x = element_text(angle = 90, size = 6.5, vjust=0.9, hjust = 0.9),
        axis.text.y = element_text(size = 6.5),
        plot.title = element_text(size = 6.5),
        legend.text = element_text(size = 6.5),
        legend.title = element_text(size = 6.5),
        strip.text.x = element_text(size = 6.5))
mofa.loading

# ggsave(plot = mofa.loading,
#        file = paste("mofa.loading", ".pdf", sep=""),
#        path = "/Users/joshgillard/Desktop/",
#        width = 18,
#        height = 6,
#        units = "cm",
#        device = "pdf")




```


### Factor regression

```{r factor regress, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 6, fig.width = 6}
dat.tmp <- mofa_fac$single_group %>%
  data.frame() %>%
  rownames_to_column("sampleID") %>%
  mutate(group = mapvalues(sampleID, from = meta.df$sampleID, to = as.character(meta.df$Severity))) %>%
  mutate(regress = case_when(group == "Control" ~ 1,
                             group == "Mild" ~ 2,
                             group == "Moderate" ~ 3,
                             group == "Severe" ~ 4)) %>%
  dplyr::select(-sampleID, -group)


mod.tmp <- lm(regress ~ . , data = dat.tmp)
summ.tmp <- summary(mod.tmp)





factor.regression <- summ.tmp$coefficients %>%
  cbind(.,confint(mod.tmp)) %>%
  data.frame() %>%
  rownames_to_column("Factor") %>%
  dplyr::rename(upperCL = 7,
                lowerCL = 6,
                p.value = 5) %>%
  dplyr::filter(Factor %like% "Factor") %>%
  mutate(Factor = factor(Factor, levels = unique(.$Factor))) %>%
  ggplot(aes(x = Factor, y = Estimate)) + 
  geom_errorbar(aes(ymin=lowerCL, ymax=upperCL), colour="black", alpha = 0.5, width = 0) +
  geom_point() +
  geom_text(aes(label = paste0("p = ", round(p.value, 3)), x = Factor, y = -0.5), size  = 2) +
  theme_minimal() +
  coord_flip() +
  labs(x = "",
       y = "Estimate",
       title = "") +
  theme(axis.title.x = element_text(size = 6.5),
        axis.title.y = element_text(size = 6.5),
        axis.text.x = element_text(angle = 0, size = 6.5, vjust=0.9, hjust = 0.9),
        axis.text.y = element_text(size = 6.5),
        plot.title = element_text(size = 6.5),
        legend.text = element_text(size = 6.5),
        legend.title = element_text(size = 6.5),
        strip.text.x = element_text(size = 6.5)) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3) +
  scale_y_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5))
factor.regression


# ggsave(plot = factor.regression,
#        file = paste("factor.regression", ".pdf", sep=""),
#        path = "/Users/joshgillard/Desktop/",
#        width = 15,
#        height = 6,
#        units = "cm",
#        device = "pdf")


```

### Factor plot

```{r factor plot colours, echo = FALSE, message = FALSE, warning  = FALSE, eval = T}
#ggplot.save <- function(...) ggplot2::ggplot(...)
#ggplot <- ggplot.save

# For setting colour scales with GGally
colfunc.red <- colorRampPalette(c("white", "#AA0114"))
ggplot <- function(...) ggplot2::ggplot(...) + scale_colour_manual(values = c("#000000",colfunc.red(4)[-1])) + scale_fill_manual(values = c("#000000",colfunc.red(4)[-1]))
unlockBinding("ggplot",parent.env(asNamespace("GGally")))
assign("ggplot",ggplot,parent.env(asNamespace("GGally")))

```


```{r factor plot, echo = FALSE, message = FALSE, warning  = FALSE, fig.height = 9, fig.width = 9}
# Visualize significant factors
mofa_fac$single_group %>% 
  data.frame() %>%
  rownames_to_column("sampleID") %>%
  mutate(group = mapvalues(sampleID, from = meta.df$sampleID, to = as.character(meta.df$Severity))) %>%
  dplyr::select(2:18, group) %>%
  ggpairs(.,
          columns = c(3,10),
          progress = TRUE,
          mapping = ggplot2::aes(colour = group),
          lower = list(continuous = my_fn),
          upper = list(continuous = my_fn),
          diag = list(discrete= "blankDiag",
                      continuous = wrap("densityDiag", alpha=0.5)),
          legend = 1
  ) +
  theme_bw() + # theme customizations
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 7),
        panel.grid.major = element_blank())

```

```{r factor plot , echo = FALSE, message = FALSE, warning  = FALSE, eval = T}
plot.factor <- mofa_fac$single_group %>% # Only factors 2 and 5
  data.frame() %>%
  rownames_to_column("sampleID") %>%
  mutate(group = mapvalues(sampleID, from = meta.df$sampleID, to = as.character(meta.df$Severity)))

p <- ggplot(plot.factor, aes(x = Factor3, y = Factor10, colour = group)) +
  geom_point(size = 2) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 6.5),
        axis.title.y = element_text(size = 6.5),
        axis.text.x = element_text(angle = 0, size = 6.5, vjust=0.9, hjust = 0.9),
        axis.text.y = element_text(size = 6.5),
        plot.title = element_text(size = 6.5),
        legend.text = element_text(size = 6.5),
        legend.title = element_blank(),
        strip.text.x = element_text(size = 6.5),
        legend.position = "bottom") #+
  #scale_colour_manual(values = c("#FF5280", "#6BE000)
factor.plot <- ggExtra::ggMarginal(p, type = "density", groupFill = TRUE)
factor.plot



# ggsave(plot = factor.plot,
#        file = paste("factor.plot", ".pdf", sep=""),
#        path = "/Users/joshgillard/Desktop/",
#        width = 9,
#        height = 9,
#        units = "cm",
#        device = "pdf")
```


# Patient characteristics analysis


```{r Cross correlate patient characteristics, echo = FALSE, message = FALSE, warning  = FALSE, eval = T}

meta.df <- read.csv("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Data/Clinical_info_June10.csv", na.strings = "NA", stringsAsFactors = T) %>%
  mutate(sampleID = as.character(sampleID)) %>% # Dummy encode severity labels for regression analysis
  mutate(Severity = case_when(Severity == "Severe" ~ 4,
                              Severity == "Moderate" ~ 3,
                              Severity == "Mild" ~ 2,
                              Severity == "Healthy" ~ 1))

meta.df <- meta.df %>%
  dplyr::rename(DSO = Days_since_symptom_onset) %>%
  mutate(Obesity = as.character(Obesity)) %>%
  mutate(Obesity = case_when(is.na(Obesity) == T ~ "NA",
                             TRUE ~ Obesity)) %>%
  mutate(Obesity = factor(Obesity, levels = c("Yes", "No", "NA")))



```




```{r Pairwise interactions, echo = FALSE, message = FALSE, warning  = FALSE, eval = F}
# Main plots
p1 <- meta.df %>%
  ggplot(aes(y = Severity, x = Age)) +
  geom_point(size = 1, alpha = 0.75, shape = 21, fill = "white") +
  geom_smooth(method = "lm", se = F, colour = "red") +
  theme_bw() + 
  scale_y_continuous(breaks=c(1, 2, 3, 4)) +
  coord_cartesian(#xlim = c(5, 20), 
                  ylim = c(0.9, 4.6)) +
  stat_cor(method = "spearman", size = 6.5 / (14/5), label.y = 4.4) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())


p2.test <- meta.df %>%
  compare_means(Severity ~ Sex, data = .) %>%
  mutate(y.position = c(4.4)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p2 <- meta.df %>%
  ggplot(aes(y = Severity, x = Sex)) +
  geom_violin(outlier.shape = NA) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0.1) +
  theme_bw() + 
  scale_y_continuous(breaks=c(1, 2, 3, 4)) +
  coord_cartesian( 
    ylim = c(0.9, 4.6)) +
  stat_pvalue_manual(p2.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p3.test <- meta.df %>%
  compare_means(Severity ~ Obesity, data = .) %>%
  mutate(y.position = c(4.2, 4.4, 4.6)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p3 <- meta.df %>%
  ggplot(aes(y = Severity, x = Obesity)) +
  geom_violin(outlier.shape = NA, width = 1.3) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0.1) +
  theme_bw() + 
  scale_y_continuous(breaks=c(1, 2, 3, 4)) +
  coord_cartesian(#xlim = c(5, 20), 
    ylim = c(0.9, 4.6)) +
  stat_pvalue_manual(p3.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p4 <- meta.df %>%
  ggplot(aes(y = Severity, x = DSO)) +
  geom_point(size = 1, alpha = 0.75, shape = 21, fill = "white") +
  geom_smooth(method = "lm", se = F, colour = "red") +
  theme_bw() + 
  scale_y_continuous(breaks=c(1, 2, 3, 4)) +
  scale_x_continuous(breaks=c(0, 20, 40, 60)) +
  coord_cartesian(xlim = c(0, 87), 
    ylim = c(0.9, 4.6)) +
  stat_cor(method = "spearman", size = 6.5 / (14/5),  label.y = 4.4) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p5.test <- meta.df %>%
  compare_means(Age ~ Sex, data = .) %>%
  mutate(y.position = c(84)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p5 <- meta.df %>%
  ggplot(aes(y = Age, x = Sex)) +
  geom_violin(outlier.shape = NA) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0) +
  theme_bw() + 
  scale_y_continuous(breaks=c(20, 40, 60, 80)) +
  coord_cartesian(
    ylim = c(15, 88)) +
  stat_pvalue_manual(p5.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p6.test <- meta.df %>%
  compare_means(Age ~ Obesity, data = .) %>%
  mutate(y.position = c(80, 84, 88)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p6 <- meta.df %>%
  ggplot(aes(y = Age, x = Obesity)) +
  geom_violin(outlier.shape = NA) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0) +
  theme_bw() + 
  scale_y_continuous(breaks=c(20, 40, 60, 80)) +
  coord_cartesian(
    ylim = c(15, 88)) +
  stat_pvalue_manual(p6.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p7 <- meta.df %>%
  group_by(Sex, Obesity) %>%
  dplyr::summarize(n = n()) %>%
  ungroup() %>%
  mutate(N_size = cut(n, breaks=c(-Inf, 10, 20, 30, 40, 50, 60, Inf), label=c(10, 20, 30, 40, 50, 60, 70))) %>%
  ggplot(aes(y = Sex, x = Obesity, size = N_size)) +
  geom_point(alpha = 1, shape = 21, fill = "white") +
  geom_text(aes(label = n), size = 6.5 / (14/5)) +
  theme_bw() + 
  scale_size_manual(values = c(1,2,3,4,5,6) * 5, guide="none") +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())


p8 <- meta.df %>%
  ggplot(aes(y = Age, x = DSO)) +
  geom_point(size = 1, alpha = 0.75, shape = 21, fill = "white") +
  geom_smooth(method = "lm", se = F, colour = "red") +
  theme_bw() + 
  stat_cor(method = "spearman", size = 6.5 / (14/5), label.y = 84) + 
  scale_y_continuous(breaks=c(20, 40, 60, 80)) +
  scale_x_continuous(breaks=c(0, 20, 40, 60)) +
  coord_cartesian(xlim = c(0, 87),
    ylim = c(15, 88)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())



p9.test <- meta.df %>%
  compare_means(DSO ~ Sex, data = .) %>%
  mutate(y.position = c(82)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p9 <- meta.df %>%
  ggplot(aes(y = DSO, x = Sex)) +
  geom_violin(outlier.shape = NA) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0) +
  theme_bw() + 
  scale_y_continuous(breaks=c(0, 20, 40, 60)) +
  coord_flip(ylim = c(0, 87)) +
  stat_pvalue_manual(p9.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

p10.test <- meta.df %>%
  compare_means(DSO ~ Obesity, data = .) %>%
  mutate(y.position = c(77, 82, 87)) %>%
  mutate(stars = cut(p, breaks=c(-Inf, 0.001, 0.01, 0.05, Inf), label=c("***", "**", "*", "ns")))

p10 <- meta.df %>%
  ggplot(aes(y = DSO, x = Obesity)) +
  geom_violin(outlier.shape = NA) +
  geom_jitter(size = 1, alpha = 0.75, shape = 21, fill = "white", width = 0.1, height = 0) +
  theme_bw() + 
  scale_y_continuous(breaks=c(0, 20, 40, 60)) +
  coord_flip(ylim = c(0, 87)) +
  stat_pvalue_manual(p10.test, label = "stars", tip.length = 0, size = 6.5 / (14/5)) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())


p1
p2
p3
p4
p5
p6
p7
p8
p9
p10

# ggsave(plot = p10,
#          file = paste("p10", ".pdf", sep=""),
#          path = "/Users/joshgillard/Desktop/",
#          width = 4,
#          height = 4,
#          units = "cm",
#          device = "pdf",
#          useDingbats = FALSE)


# Axis labels
p1.labs <- p1 + theme(axis.text.y = element_text(size = 6.5),
        axis.text.x = element_text(size = 6.5),
        axis.title = element_blank())

# ggsave(plot = p1.labs,
#          file = paste("p1.labs", ".pdf", sep=""),
#          path = "/Users/joshgillard/Desktop/",
#          width = 4,
#          height = 4,
#          units = "cm",
#          device = "pdf",
#          useDingbats = FALSE)

p5.labs <- p5 + theme(axis.text.y = element_text(size = 6.5),
        axis.text.x = element_text(size = 6.5),
        axis.title = element_blank(),
        axis.ticks = element_line())

# ggsave(plot = p5.labs,
#          file = paste("p5.labs", ".pdf", sep=""),
#          path = "/Users/joshgillard/Desktop/",
#          width = 4,
#          height = 4,
#          units = "cm",
#          device = "pdf",
#          useDingbats = FALSE)


p7.labs <- p7 + theme(axis.text.y = element_text(size = 6.5),
        axis.text.x = element_text(size = 6.5),
        axis.title = element_blank(),
        axis.ticks = element_line())

# ggsave(plot = p7.labs,
#          file = paste("p7.labs", ".pdf", sep=""),
#          path = "/Users/joshgillard/Desktop/",
#          width = 4,
#          height = 4,
#          units = "cm",
#          device = "pdf",
#          useDingbats = FALSE)

p10.labs <- p10 + theme(axis.text.y = element_text(size = 6.5),
        axis.text.x = element_text(size = 6.5),
        axis.title = element_blank(),
        axis.ticks = element_line())

# ggsave(plot = p10.labs,
#          file = paste("p10.labs", ".pdf", sep=""),
#          path = "/Users/joshgillard/Desktop/",
#          width = 4,
#          height = 4,
#          units = "cm",
#          device = "pdf",
#          useDingbats = FALSE)


```

```{r covariates corr, echo = FALSE, message = FALSE, warning  = FALSE, eval = T}

ethnicity.df <- read.csv("/Users/jgillard/Desktop/Bioinformatics/Multiomic COVID19/Multiomics_covariates/Ethnicity.csv", na.strings = "NA", stringsAsFactors = T) %>%
  mutate(sampleID = as.character(sampleID)) %>% # Dummy encode severity labels for regression analysis
  mutate(Severity = case_when(Severity == "Severe" ~ 4,
                              Severity == "Moderate" ~ 3,
                              Severity == "Mild" ~ 2,
                              Severity == "Healthy" ~ 1)) %>%
  mutate(Hispanic = case_when(EthnicityHispanicLatino == "No" ~ 0,
                              EthnicityHispanicLatino == "Yes" ~ 1)) %>%
  dplyr::select(sampleID, Hispanic)


meta.df <- meta.df %>%
  full_join(., ethnicity.df, by = "sampleID") %>%
  mutate(Sex = case_when(Sex == "M" ~ 0,
                         Sex == "F" ~ 1)) %>%
  mutate(Obesity = case_when(Obesity == "No" ~ 0,
                             Obesity == "Yes" ~ 1)) %>%
  dplyr::select(sampleID, Severity, Age, Sex, Hispanic, Obesity, DSO)
  


coef.mat <- meta.df %>%
  dplyr::select(-sampleID) %>%
  cor(., method = "pearson", use = "pairwise.complete.obs")
p.mat <- meta.df %>%
  dplyr::select(-sampleID) %>%
  cor.mtest(., method = "pearson", use = "pairwise.complete.obs")


corrplot(coef.mat, p.mat = p.mat , 
         method = 'square', type = 'lower', insig='blank', 
         number.cex = 0, 
         order = 'AOE', 
         diag=FALSE, 
         col = rev(brewer.pal(n = 10, name = 'RdBu')),
         tl.col = 'black')

```






